\chapter{Representation}
\label{chap:represenation}

% NOTE: taken from KnowRob2.0
% One of the most powerful components of the human memory system is the
% episodic memory. 
When somebody talks about the deciding goal in the
last soccer world championship many of us can ``replay'' the episode
in our ``mind's eye''. %, or if somebody talks about a cross court hit
%in tennis we can recall how such a hit feels in the arm.
The memory
mechanism that allows us to recall these very detailed pieces of
information from abstract descriptions is our episodic memory.
Episodic memory is powerful because it allows us to remember special
experiences we had. It can also serve as a ``repository'' from which
we learn general knowledge.

% NOTE: taken from KnowRob2.0
EASE integrates episodic memories
deeply into the knowledge acquisition, representation, and processing
system. Whenever a robotic agent performs, observes, prospects, and
reads about an activity, it creates an episodic memory. An episodic
memory is best understood as a video that the agent makes of the
ongoing activity 
%in its inner world knowledge base
coupled with a very
detailed story about the actions, motions, their purposes, effects,
the behavior they generate, the images that are captured, etc.

% % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % %
\section{Upper-Level Model}
\label{sec:narrative:commitments}

% NOTE: taken from SOMA paper
We decided to base our model on the DOLCE+DnS Ultralite (DUL) foundational framework~\cite{DOLCE2003}.
This decision is greatly motivated by their underlying ontological commitments.
%The decision to base our model on the DOLCE+DnS Ultralite foundational framework, is greatly motivated by their underlying ontological commitments.
Firstly, DUL is not a revisionary model, but seeks to express stands that shape human cognition. Furthermore it assumes a reductionist approach -- rather than capturing, for example, the flexibility of our usage of objects via multiple inheritance in a multiplicative manner, we commit to a reduced {\it ground} classification and use a {\it descriptive} approach for handling this flexibility. For this a primary branch of the ontology represents the ground {\bf physical model}, e.g. objects and actions, while a secondary branch represents the {\bf social model}, e.g. roles and tasks. All entities in the social branch would not exist without humans, i.e. they constitute social objects that represent concepts about or descriptions of ground elements. 

% NOTE: taken from SOMA paper
Every axiomatization in the physical branch can, therefore, be regarded as expressing some physical context whereas axiomatizations in the descriptive social branch are used to express social contexts. A set of dedicated relations is provided that connect both branches. For example, as detailed in Section~\ref{subsec:roles}, the relation \emph{classifies} connects ground objects, e.g. a hammer, with the roles they can play, i.e. potential classifications. Thus, we can state that a hammer can in some context be conceptualized as a murder weapon, a paper weight or a door stopper. Nevertheless, neither its ground ontological classification as a tool will change nor will hammers be subsumed as kinds of door stoppers, paper weights or weapons via multiple inheritance. Following a quick overview of the central modules of SOMA where these commitments apply, we will provide detailed examples of where and how our commitments apply in Sections~\ref{sec:perdurants} and~\ref{sec:endurants}.

% % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % %
\input{content/representation/neem-narrative/neem-narrative}

% % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % %
\input{content/representation/neem-experience/neem-experience}

% % % % % % % % % % % % % % % % %
% % % % % % % % % % % % % % % % %
\section{Extending the Representation}
\label{sec:extension}
\subsubsection{New Tasks Definition}
\label{sec:extension,subsec:task}
\todo{Text is copied from the robot-neem section.}
A task might be something like cutting, stopping or accelerating.
To be able to semantically log the task, you will need first define the task in the \easeAct.
Make sure that the new action is a child of the \textit{CommunicationTask}, \textit{MentalTask} or \textit{PhysicalTask}.
