\chapter{NEEM-Experience}

\neemexp captures low-level information about experienced activities
represented as time series data streams.
This data has often no or only unfeasible
lossless representation as facts in a knowledge base.
To make this data \emph{knowledgable}, procedural hooks
are defined in the ontology to compute symbols from the experience data,
and to embed these symbols in logic-based reasoning.

The data is stored in a NoSQL database using JSON documents.
Each individual type of data is stored in a separate collection
named according to the type of data stored in the collection.
When imported, the knowledge system stores the data in a
\mongodb\footnote{https://www.mongodb.com/} server, for which
the knowledge system implements a client for querying the data
during question answering.
The query cursor concept employed in \mongodb integrates
nicely with backtracking based search employed in the knowledge system.
It further scales well to large amount of data and can be distributed amongst
clusters through built-in automatic sharding.

The data in \neemexps is represented as time series
and indexed in time order.
The different experience data types need to define a dedicated
time key for computing the search index.

% the \neemexp consists only of \tf data \cite{tf}.
% When executing an experiment in projection, we are collecting \tf data from the map, objects and the robot.
% In real world experiments we log only the \tf data from the map and the robotic agent.
% We are storing the \tf data in a \mongodb\footnote{https://www.mongodb.com/}.
% In the next iteration of the document we will provide links where you can download the required tools for log \tf data for NEEMs.

The experience data in \neems has individual characteristics regarding
the format, compressed representation, and what symbols the 
knowledge system can abstract from the data.
In this chapter, we describe these aspects for the experience data types covered 
in \neem version \neemversion.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Pose Data}
\input{content/neem-experience/pose-data}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Image Data}
% \input{content/neem-experience/image-data}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{Gaze Data}
% \input{content/neem-experience/gaze-data}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section{EEG Data}
% \input{content/neem-experience/eeg-data}
